### 2025-02-26 manuscript/manuscript.qmd
text edits to results and conclusions

#### Outlier analysis

@@

Figure 4 shows forest and funnel plot visualizations from the random-effects model of the best-performing models in classification studies. _MCC_s vary across a wide range, reporting _MCC_s between 0.55 to 0.98. Table 4 indicates that using the best model from each study increases performance relative to all models ($MCC$ = 0.8684 95%CI: 0.7475-0.9336), yet slightly increases heterogeneity ($\tau^2$ = 0.3180, $I^2$ = 99.8%).

@@

Heterogeneity issues

@@

The number of features analyzed does not significantly impact results (_Q_(2) = 3.91, _p_ = 0.1419), though $MCC$ shows differences on averages based on the number of features (< 30 F: $MCC$ = 0.9293; 30-300 F: $MCC$ = 0.8458; >300 F: $MCC$ = 0.7754). Conversely, model classes differ significantly (_Q_(3) = 4.22, _p_ = 0.239). Specifically, neural networks outperform all other model types ($MCC$ = 0.9307), followed by SVMs ($MCC$ = 0.870) and tree-based methods ($MCC$ = 0.854), then linear methods ($MCC$ = 0.728). Studies did not differ based on whether stimuli comprised one or multiple genres (_Q_(1) = 0.12, _p_ = 0.7324), or whether studies employed binary or multi-class models (_Q_(1) = 0.03, _p_ = 0.8694).

Brief summary

<!--
Notes to results: In the pre-registration, we promised the following:

- groupings based on genre, model complexity (high, medium, low), model validation (exists or not)
- we promised to carry out sensitivity analyses based on type of journal the studies were published in.

-->

@@

We initially identified 96 studies involving MER models, but in many cases the reporting standards or other quality issues or the lack of dedicated focus on music led us to include 34 studies in the analysis that overall reported 290 models, 204 related to regression and 86 classification. Comparing the most accurate model in each study, we observed reasonably accurate prediction of valence (_r_ = 0.67 [0.56, 0.76]), and arousal (_r_ = 0.81 [0.74, 0.86]) in regression studies. For both affective dimensions of the circumplex model, linear methods (valence _r_ = 0.78, arousal _r_ = 0.88) and tree-based models (valence _r_ = 0.75, arousal _r_ = 0.81) outperformed neural networks (valence = 0.47, arousal = 0.66) and support vector machines (SVMs) (valence _r_ = 0.58, arousal _r_ = 0.79). In contrast, neural networks performed most accurately in classification studies (_MCC_ = 0.93), followed by SVMs (_MCC_ = 0.87) and tree-based models (_MCC_ = 0.85). The models applied to a variety of datasets differ substantially in terms of size, annotation principles, and the number of features and sources of these features.

@@

The results of the meta-analysis offer several important insights. First, compared to a 2013 report by @barthet2013, the glass ceiling for valence has risen from _r_ = 0.51 to 0.67. Conversely, predictive accuracy for arousal has shown no improvement (reported already in 2013 at _r_ = 0.83). For classification, studies reached variable classification rates from 0.497 using _F~1~_ score [@sanden2011empirical] to 79% average precision [@trohidis2008multi] in the past, whereas now the accuracy is generally notably better, here _MCC_=0.87, albeit the differences in metrics.

@@

Future reports should contain viable information about the data, models, and success metrics. The modelling process should include a description of cross-validation, feature extraction technique, feature reduction (if used), and actual accuracy measures. We would recommend future studies to use Matthews correlation coefficient (_MCC_) for classification studies as currently there are numerous metrics (_F~1~_, precision, recall, balanced accuracy, etc.), which do not operate in the same range. The situation is regression studies tends to be less volatile, but model accuracy may be reported with error measures (Mean Squared Error, Root Mean Squared Error, or Mean Absolute Error, or MSE, RMSE, MAE), which cannot be converted to relative measure such as _R^2_ without knowing the variance of the data.

@@

>>>>

#### Study heterogeneity

@@

Figure 4 shows forest and funnel plot visualizations from the random-effects model of the best-performing models in classification studies. $MCC$s vary across a wide range, reporting $MCC$s between 0.55 to 0.98. Table 4 indicates that using the best model from each study increases performance relative to all models ($MCC$ = 0.8684 95%CI: 0.7475-0.9336), yet slightly increases heterogeneity ($\tau^2$ = 0.3180, $I^2$ = 99.8%).

@@

More heterogeneity issues?

@@

The number of features (as classified into under 30, between 30 and 300, and above 300 features) analyzed does not significantly impact results (_Q_(2) = 3.91, _p_ = 0.1419), though $MCC$ shows differences on averages based on the number of features (< 30 F: $MCC$ = 0.9293; 30-300 F: $MCC$ = 0.8458; >300 F: $MCC$ = 0.7754). Conversely, model classes differ significantly (_Q_(3) = 4.22, _p_ = 0.239). Specifically, neural networks outperform all other model types ($MCC$ = 0.9307), followed by SVMs ($MCC$ = 0.870) and tree-based methods ($MCC$ = 0.854), then linear methods ($MCC$ = 0.728). Studies did not differ based on whether stimuli comprised one or multiple genres (_Q_(1) = 0.12, _p_ = 0.7324), or whether studies employed binary or multi-class models (_Q_(1) = 0.03, _p_ = 0.8694).

NOTE. Model classes do not seem to differ significantly!?

<!--
Notes to results: In the pre-registration, we promised the following:
- groupings based on genre, model complexity (high, medium, low), model validation (exists or not)
- we promised to carry out sensitivity analyses based on type of journal the studies were published in.
-->

Anything missing?

- Leave-one-out sensitivity analyses?

@@

We initially identified 96 studies involving MER models, but in many cases the reporting standards or other quality issues or the lack of dedicated focus on music led us to include 34 studies in the analysis that overall reported 290 models, 204 related to regression and 86 classification. Comparing the most accurate model in each study, we observed reasonably accurate prediction of valence (_r_ = 0.67 [0.56, 0.76]), and arousal (_r_ = 0.81 [0.74, 0.86]) in regression studies. For both affective dimensions of the circumplex model, linear methods (valence _r_ = 0.78, arousal _r_ = 0.88) and tree-based models (valence _r_ = 0.75, arousal _r_ = 0.81) outperformed neural networks (valence = 0.47, arousal = 0.66) and support vector machines (SVMs) (valence _r_ = 0.58, arousal _r_ = 0.79). In contrast, neural networks performed most accurately in classification studies ($MCC$ = 0.93), followed by SVMs ($MCC$ = 0.87) and tree-based models ($MCC$ = 0.85). The models applied to a variety of datasets differ substantially in terms of size, annotation principles, and the number of features and sources of these features.

@@

The results of the meta-analysis offer several important insights. First, compared to a 2013 report by @barthet2013, the glass ceiling for valence has risen from _r_ = 0.51 to 0.67. Conversely, predictive accuracy for arousal has shown no improvement (reported already in 2013 at _r_ = 0.83). For classification, studies reached variable classification rates from 0.497 using _F~1~_ score [@sanden2011empirical] to 79% average precision [@trohidis2008multi] in the past, whereas now the accuracy is generally notably better, here $MCC$ = 0.87, albeit the differences in metrics.

@@

Future reports should contain viable information about the data, models, and success metrics. The modelling process should include a description of cross-validation, feature extraction technique, feature reduction (if used), and actual accuracy measures. We would recommend future studies to use Matthews correlation coefficient ($MCC$) for classification studies as currently there are numerous metrics (_F~1~_, precision, recall, balanced accuracy, etc.), which do not operate in the same range. The situation is regression studies tends to be less volatile, but model accuracy may be reported with error measures (Mean Squared Error, Root Mean Squared Error, or Mean Absolute Error, or $MSE$, $RMSE$, $MAE$), which cannot be converted to relative measure such as $R^2$ without knowing the variance of the data.

@@

## Implications and Conclusion
